Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-04-06 paper.tex
Add proposed table from Jackson et al, commented out
Add in some rough notes for pango lineage analyses

@@

In this section we examine the Long ARG to determine how well it reflects
the structure of previously identified recombinant lineages.
There are [high level summary of the pango recombinant lineages and samples
in the Long ARG].
long ARG (subsampled to mid-2022) contains samples
from 33 Pango-designated recombinant lineages (XA to XAK; Supplementary Table)
[Probably could have a table here to summarise?]

@@

@@

>>>>

% jk: this is the proposal from issue #33

% \begin{tabular}{ll|clll}
% \toprule
% A Group (XA) & Jackson        & 21,255–21,613 & B.1.177/B.1.1.7 \\
%              & \texttt{sc2ts} &  21,256--22,227 & B.1.177.18/B.1.1.7 \\
% \midrule
% B Group & Jackson        &  6,528--6,954 & B.1.36/B.1.1.7  \\
%         & \texttt{sc2ts} &  6,529--6,954 & B.1.36/B.1.1.7  \\
% \midrule
% C Group & Jackson        &  25,997--27,972 &  B.1.1.7/B.1.221 \\
%         & \texttt{sc2ts} &  25,996--27,441 &  B.1.1.7/B.1.221 \\
% \midrule
% D Group & Jackson        &  21,575--23,063 &  B.1.36.17/B.1.1.7 \\
%         & \texttt{sc2ts} &  22,445--23,063 &  B.1.36.39/B.1.1.7 \\
% \midrule
% CAMC-CBA018 & Jackson        &  20,389--21,254 & B.1.177/B.1.1.7 \\
%             & \texttt{sc2ts} & 16,177--21,255 & B.1.177/B.1.1.7 \\
% \midrule
% \end{tabular}

@@

In this section we examine the Long ARG to determine how well it reflects
the structure of previously identified recombinant lineages.
There are [high level summary of the pango recombinant lineages and samples
in the Long ARG].
long ARG (subsampled to mid-2022) contains 749 samples
from 33 Pango-designated recombinant lineages (XA to XAK; Supplementary Table)
[Probably could have a table here to summarise?]

% \item Total of 2078 recombinants.
% \item 749 samples from 33 Pango X lineage (recombinant) designations
% \item \alert{All but} 2 XP samples inferred to have recombinant ancestry
% (Almost certainly missing XP because of our alignment masking approach)

All but 2 XP samples inferred to have recombinant ancestry.

@@

[TODO need to be clear if we're looking at single monophyletic origins
here, or whether we include complicated things like XAG etc or not.
Probably not?]

% \item 14 out of 33 of the Pango X lineage inferred to have a single
% recombinant origin. (XA, XAA, XAC, XAE, XAG, XAK, XF, XG, XH, XL,
%  XR, XT, XW, XY)
% \item Total of 148 samples (max 32, average of 10 per group)
% \item XAA and XAG share the same origin, all others are distinct
%     \begin{itemize}
%         \item Ours: BA.1 \& BA.2.9 (6472 -- 8393)
%         \item XAA: BA.1* \& BA.2* (8393 -- 9343)
%         \item XAG: BA.2* \& BA.1* (6516 -- 8392)
%         \item At least one XAB lineage (50 samples, with 10 origins)
%         also descends from this recombinant---unclear what's happening
%     \end{itemize}
% \end{itemize}
% \end{block}
% \end{frame}

@@

% \begin{frame}
% \begin{block}{Multiple origin Pango X lineages}
% \begin{itemize}
% \item For the remaining 18, the story is more complicated.
% \item XJ, XU and XV each have 3 samples and 3 recombinants
% \item One recombinant shared among 5 lineages (XAA, XAB, XAG, XQ, XU)
% \item XE has 170 samples and 12 recombinants; some are very similar
% (Official: BA.1* \& BA.2, 10448--11287):
% \\
% \vspace{1em}
% \begin{tabular}{llll}
% \toprule
% Left parent & Right parent & \multicolumn{2}{l}{Breakpoint interval}\\
% \midrule
% BA.1.15 & BA.2  &9867  &10198\\
% BA.1.15 & BA.2  &9867  &10198\\
% BA.1.1  &BA.2  &10451  &11537\\
% BA.1.1.5 & BA.2 &10448 & 11537\\
% BA.1.15 & BA.2  &10448  &12880\\
% BA.1.17 & BA.2  &10448 & 11537\\
% BA.1  &BA.2.9  &10448 & 11537\\
% BA.1  &BA.2  &12764  &13195\\
% \bottomrule
% \end{tabular}
% % BA.3  &BA.1.17  &22674  &22674\\
% % BA.1.1.15 Recombinant 10448 11537
% % BA.1.1 Recombinant 10030 10447
% % Recombinant BA.2 17411 19955
% \end{itemize}
% \end{block}
% \end{frame}

\begin{figure}
\begin{tabular}{cc}

Example lineage subgraph
 &

\begin{tabular}{llll}
\toprule
Left parent & Right parent & \multicolumn{2}{l}{Breakpoint interval}\\
\midrule
BA.1.15 & BA.2  &9867  &10198\\
BA.1.15 & BA.2  &9867  &10198\\
BA.1.1  &BA.2  &10451  &11537\\
BA.1.1.5 & BA.2 &10448 & 11537\\
BA.1.15 & BA.2  &10448  &12880\\
BA.1.17 & BA.2  &10448 & 11537\\
BA.1  &BA.2.9  &10448 & 11537\\
BA.1  &BA.2  &12764  &13195\\
\bottomrule
\end{tabular}
% BA.3  &BA.1.17  &22674  &22674\\
% BA.1.1.15 Recombinant 10448 11537
% BA.1.1 Recombinant 10030 10447
% Recombinant BA.2 17411 19955
\end{tabular}
\caption{A. Example subgraph for a multi origin example. B. details
of inferred breakpoints for different recombination events inferred
for XE samples (170 samples and 12 inferred recombination nodes; some are very
similar (Official: BA.1* \& BA.2, 10448–11287)}
\end{figure}
